---
title: "Generating Templates with table_templates.R"
author: "[Your Name]"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Table Templates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
This vignette demonstrates how to use pre-defined templates to create new R scripts for generating tables.The core idea is to provide a set of standard templates for common statistical tables, such as those for demography (`t_dm_slide`). Users can thenmodify them to suit their specific datasets and analysis needs.

## List Available Templates
Before generating a template, you can view all supported table templates using the ` list_all_templates()` function:


```{r, include = TRUE, eval = FALSE}
library(autoslider.core)
list_all_templates()

```
This will return a character vector of supported templates (e.g., "t_dm_slide", "t_ae_slide", etc.), each corresponding to a table type stored in the `R/` folder of the package.

## Generate a Table Template
Use the `use_template()` function to generate a template script that you can customize. For example:

```{r_function_call, include = TRUE, eval = FALSE}
use_template(template = "t_dm_slide", function_name = "demography")
```
This will create a file with the demography template at `./programs/R/demography.R`

### Notes
- The `template` argument must match one of the entries in list_all_templates(). 

- If a file with the same name already exists, it will not be overwritten unless you set `overwrite = TRUE`.

- In interactive mode, the generated file will open in your R editor for immediate editing.


## Customize the Template
Once the script is generated, you can modify it to suit your dataset or statistical needs. For instance, you can:


- Add or remove table columns

- Customize labels and formatting

- Modify the summary statistics



