---
title: "Generating AI-Powered speaker notes"
date: "`r Sys.Date()`"
output:
    rmarkdown::html_document:
        theme: "spacelab"
        highlight: "kate"
        toc: true
        toc_float: true
author: 
  - Yolanda Zhou ([`yolandazzz13`])

vignette: >
  %\VignetteIndexEntry{Deploy LLM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
    markdown:
        wrap: 72
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Overview
This guide explains how to use the `get_ai_notes()` function to automatically add analytical annotations and summaries to your slides using AI tool.

## Prerequisite

Before you begin, ensure you have the following ready:

- A Prompt File (`prompt.yml`) containing the instructions for the AI model.

- Access to a local LLM instance. 

Please see `autoslideR.Rmd` for some readily available spec and filter examples and detailed instructions on how to use them. If you do not have an LLM model installed yet, see appendix on how to deploy your own local LLM.

## The Workflow
The process involves three main steps: generating the initial slide data, adding the AI footnotes, and then creating the final PowerPoint file.

### Step 1: Initial Slide Generation
First, generate the core slide outputs from your specification file. This process reads your spec, filters it for the desired programs, and generates the basic table and plot objects.

```{r, eval = FALSE}
spec <- read_spec("path/to/your/spec.yml")
filters::load_filters("path/to/your/filter.yml")
prompt_list <- get_prompt_list("path/to/your/prompt.yml")

# Generate the initial outputs
outputs <- spec %>%
  filter_spec(program %in% c("t_dm_slide")) %>%
  generate_outputs(datasets = my_datasets) %>%
  decorate_outputs()
```


### Step 2: Adding AI Footnotes
Next, pass the outputs object to the `get_ai_notes()` function. This function iterates through your outputs, and for each one that has a corresponding prompt, it communicates with the specified LLM model to generate response in the form of a speaker note.

You can configure the function to point to different AI platforms.

#### Example: Using a Local LLM 

This is ideal for local development. It assumes you have Ollama on your local machine and is developing in a Docker container.

```{r, eval = FALSE}
outputs_ai <- get_ai_notes(
  outputs = outputs,
  prompt_list = prompt_list,
  platform = "ollama",
  base_url = "http://host.docker.internal:11434", # URL for Ollama if R in a Docker container
  model = "deepseek-r1:1.5b" # The LLM model name
)
```
The parameters you pass in might depend on your specific situation. In general:

- `platform`: Set to `ollama`.

- `base_url`: Points to where your `Ollama` instance is running.

- `model`: The name of a model you have pulled in `Ollama.`


### Step 3: Create the Final PowerPoint File
Finally, take the modified `outputs_ai` object and pass it to `generate_slides()` to create the presentation. The footnotes will be automatically included on the relevant slides.


```{r, eval = FALSE}
outputs_ai %>%
  generate_slides(outfile = "My_AI_Presentation.pptx")
```

This will produce a PowerPoint file with your tables and plots, and enhanced with AI generated analysis.


## Appendix
### Running Your Local LLM

First of all, we need to download the Ollama tool at https://ollama.com/download. Once Ollama is installed, you can run any model from its library with a single command. For this example, we'll use deepseek-r1:1.5b.

Open your command-line tool (e.g., Terminal, Windows PowerShell).

Type the following command and press Enter. Ollama will automatically download the model, which may take a few minutes.

```{bash, eval=FALSE}
$ ollama run deepseek-r1:1.5b
```

After the process completes, you will see a success message and a new prompt, like this:
```{bash, eval=FALSE}
$ >>> Send a message (/? for help)
```
This means you have successfully installed and are now running a local LLM. Feel free to start a conversation and play around with some prompts!

